Clase N1

### 2024-08-31 ejercicios/linear_models_panel_data.qmd
Replication of Figures and Tables 2-3.

# Modelos de datos panel
```{r}
rm(list=ls(all=TRUE))
# Cargar paquetes necesarios
library(dplyr); library(haven); library(plm); library(tidyverse); library(summarytools); library(pder);
library(ggplot2);library(tidyr);library(panel)
```

@@

# Renombrar variables
data <- SeatBelt %>%
  rename(
    fatalities_occ = farsocc,
    fatalities_non_occ = farsnocc,
    belt_usage = usage,
    income = percapin,
    unemployment = unemp,
    age = meanage,
    pct_black = precentb,
    pct_hispanic = precenth,
    density_urban = densurb,
    density_rural = densrur,
    violent_crime = viopcap,
    property_crime = proppcap,
    vmt_rural = vmtrural,
    vmt_urban = vmturban,
    fuel_tax = fueltax,
    speed_65 = lim65,
    speed_70plus = lim70p,
    drinking_age_21 = mlda21,
    bac_limit_08 = bac08,
    belt_law_secondary = ds,
    belt_law_primary = dp,
    belt_law_primary_after_secondary = dsp
  )
```
```{r}

@@

# Dando estructura panel a la data
data <-  pdata.frame(data, c("state","year"))
attach(data)
```
```{r}
# Pool OLS
pool <- plm(fatalities_occ ~ age + belt_usage + violent_crime + unemployment + income +pct_black+pct_hispanic , data = data, model = "pooling", effect= "twoway")

@@

# Breusch-Pagan test: the LM test helps you decide between a random effects regression and a pooled OLS regression.
# The null hypothesis in the LM test is that variances across entities is zero. This is, no
# significant difference across units (i.e. no panel effect)
plmtest(pool, type="bp")
```
```{r}
fe <- plm(fatalities_occ ~ age + belt_usage + violent_crime + unemployment + income +pct_black+pct_hispanic, 
          data = data, model = "within", effect = "twoway")
summary(fe)

>>>>

# Modelos de datos panel

```{r}
rm(list=ls(all=TRUE))
# Cargar paquetes necesarios
library(dplyr); library(haven); library(plm); library(tidyverse); library(summarytools); library(pder);
library(ggplot2);library(tidyr);library(panel)
```

@@

# Renombrar variables
data <- SeatBelt %>%
  rename(
    fatalities_occ = farsocc,
    fatalities_non_occ = farsnocc,
    belt_usage = usage,
    income = percapin,
    unemployment = unemp,
    age = meanage,
    pct_black = precentb,
    pct_hispanic = precenth,
    density_urban = densurb,
    density_rural = densrur,
    violent_crime = viopcap,
    property_crime = proppcap,
    vmt_rural = vmtrural,
    vmt_urban = vmturban,
    fuel_tax = fueltax,
    speed_65 = lim65,
    speed_70plus = lim70p,
    drinking_age_21 = mlda21,
    bac_limit_08 = bac08,
    belt_law_secondary = ds,
    belt_law_primary = dp,
    belt_law_primary_after_secondary = dsp
  )
```

@@

# Dando estructura panel a la data
data <-  pdata.frame(data, c("state","year"))
attach(data)
```

```{r}
# Pool OLS
pool <- plm(fatalities_occ ~ age + belt_usage + violent_crime + unemployment + income +pct_black+pct_hispanic , data = data, model = "pooling", effect= "twoway")

@@

# Breusch-Pagan test: the LM test helps you decide between a random effects regression and a pooled OLS regression.
# The null hypothesis in the LM test is that variances across entities is zero. This is, no
# significant difference across units (i.e. no panel effect)
plmtest(pool, type="bp")
```

```{r}
fe <- plm(fatalities_occ ~ age + belt_usage + violent_crime + unemployment + income +pct_black+pct_hispanic, 
          data = data, model = "within", effect = "twoway")
summary(fe)
